---
title: "FA Assignment-11915050"
output:
  word_document: default
  html_document:
    df_print: paged
  pdf_document: default
---

```{r}
# setup chunk

suppressPackageStartupMessages({
  
  if (!require(forecast)){install.packages("forecast")}
  if (!require(zoo)){install.packages("zoo")}
  if (!require(fpp2)){install.packages("fpp2")}
  if (!require(ggpubr)){install.packages("ggpubr")}
  if (!require(tseries)){install.packages("tseries")}
  library(forecast) 
  library(zoo)
  library(fpp2)
  library(ggpubr)
  library(tseries)
})
```

## loading the Auscafe dataset and transformation to a timeseries
```{r pressure, echo=FALSE}
Auscafe.data <- read.csv("D:\\FA\\auscafe.csv")
head(Auscafe.data)
tail(Auscafe.data)

Auscafe.ts <- ts(Auscafe.data$value, start = c(1982,4), end = c(2017, 09), freq = 12)
plot(Auscafe.ts, xlab = "Time", ylab = "Aus cafe")
```

## Decomposing the timeseries into components using additive and multiplicative model
```{r}
Auscafe1 <- decompose(Auscafe.ts, type= "additive")
g1 <- autoplot(Auscafe1)
Auscafe2 <- decompose(Auscafe.ts, type= "multiplicative")
g2 <- autoplot(Auscafe2)
ggarrange(g1,g2)
```
# A clear trend and seasonality is visible in the plots
# (a) Partition the data keeping last 24 months data as validation set using R.
```{r}
stepsAhead <- 24
length(Auscafe.ts)
# Length of training data set = length of series - validation period (24)
nTrain <- length(Auscafe.ts) - stepsAhead
train.ts <- window(Auscafe.ts, start = c(1982, 4), end = c(1982,  nTrain + 4 - 1))
length(train.ts)
valid.ts <- window(Auscafe.ts, start = c(1982, nTrain + 4), end = c(1982, nTrain + stepsAhead + 4 -1))
length(valid.ts)

# fitting a polynomial trend line on training data-set and forecasting it for validation dataset.
Auscafe.lm <-  tslm(train.ts ~ poly(trend, 2))
Auscafe.lm.pred <- forecast(Auscafe.lm, h = stepsAhead, level = 0)

# Plotting the trend-lines and fitted data
plot(Auscafe.lm.pred,  ylab = "Auscafe", xlab = "Time", bty = "l", xaxt = "n", xlim = c(1982,2020), ylim = c(0.5, 5.5), main = "", flty = 2)
axis(1, at = seq(1982, 2020, 1), labels = format(seq(1982, 2020, 1))) 
lines(Auscafe.lm$fitted, lwd = 2)
lines(valid.ts)
lines(c(2017.75 - 2, 2017.75 - 2), c(0, 4), col= "blue") 
lines(c(2017.75, 2017.75), c(0, 4), col = "red")
text(2004.5, 3.5, "Training", col = "Blue")
text(2016.5, 2.5, "Validation", col = "Red")
text(2019.5, 3.5, "Future", col = "Orange")
arrows(2017.75 - 2, 3.2, 1984, 3.2, code = 3, length = 0.1, lwd = 1,angle = 20)
arrows(2017.75, 3, 2017.75 - 2, 3, code = 3, length = 0.1, lwd = 1,angle = 20)
arrows(2018, 3.2, 2020, 3.2, code = 3, length = 0.1, lwd = 1, angle = 20)
```

#(b) Draw the time plot, ACF, PACF for the above training data
```{r}
autoplot(train.ts)
Acf(train.ts, lag.max = 20, main = "")
Pacf(train.ts, lag.max = 20, main = "")

# PACF plot indicates that lag-1 has the major spike containing the auot-correlationfactor AR-1
```
#(c) Fit an exponential trend model with multiplicative seasonality on the training set#. Are the residuals correlated.

```{r}
# Observing the seasonal component and plots before fitting a seasonal model
autoplot(decompose(train.ts))
# Since there is seasonal component we need to first deseasonalize the data 
# fit a exponential trend model and then re-seasonalize the data.

# Deseasonalizing
train.noseason <- train.ts - decompose(train.ts)$seasonal
plot(train.noseason)

# Fitting an exponenetial model to train dataset by using tslm(timeseries linear regression fit model)
# lambda = 0 applies box-cox transformation
# log(yt) = Beta0 + Beta1 * t + error
train.exp.trend <- tslm(train.noseason ~ trend, lambda = 0)  
train.exp.pred <- forecast(train.exp.trend, h=length(valid.ts), level = 0) 

summary(train.exp.trend)


```
##(d) Use the fitted model to forecast the time series on the validation data. Plot the actual vs fitted and residual plot and estimate the RMSE on the validation set.
```{r}
# training fitted exponential model over full dataset
Trendmodel <-  ts(union(train.exp.trend$fitted.values,train.exp.pred$mean),start=c(1982,1),frequency = 12) 
seasonality <- decompose(Auscafe.ts)$seasonal
## Adding back the seasonality component to get the forecast
forecast <- Trendmodel + seasonality

plot(Auscafe.ts, col="blue", lwd=2)
lines(forecast,col="brown", lwd=2)
legend(1995,4, legend=c("Actual", "Forecast"),  col=c("blue", "brown"), lty=1:1)
```

```{r}

plot(train.exp.trend$residuals, main= "Residual Plot", ylab="Forecast Error",lwd=2,xlim = c(1982,2020))
lines(valid.ts-train.exp.pred$mean, col="blue",lwd=2)
abline(v=2015.75, col="red", lwd=2)
text(2012,-0.2, "Training ")
text(2018,-0.25, "Validation")

# Residuals have auto co-relations.

```


3. Refer to the housing sales dataset, hsales2.csv from January 1997 through November 1995.(Use the total dataset, no need to partition)
(a) Draw the time plot, ACF, PACF for the above data using R.

```{r}
hsales.data <- read.csv("D:\\FA\\hsales2.csv")
hsales.ts <- ts(hsales.data$hsales, start = c(1987,1), end = c(1995, 11), freq = 12)
plot(hsales.ts, xlab = "Time", ylab = "House sales")

autoplot(decompose(hsales.ts))
decompose(hsales.ts)$seasonal
decompose(hsales.ts)$trend
decompose(hsales.ts)$random


Acf(hsales.ts, lag.max = 20, main = "")
Pacf(hsales.ts, lag.max = 20, main = "")

# Observation 
# ACF plot has sinusoidal pattern
# PACF plot has a significant spike at lag-1, a comparatively  insignificant spikes at 9 and 10.
```

b) Use differencing method to decompose the data into trend, seasonality and noise. Test Whether the noise has correlation left using both graphical approach and statistical test.

```{r}
ndiffs(hsales.ts) ## Number of differences required to make it stationary

kpss.test(hsales.ts) ## Null Hypothesis- stationary Time series
```
## kpss test Null Hypothesis states that its a stationary time series
## As a result of the test KPSS level = 0.25174 - fail to reject null
## this series is a stationary series


```{r}
par(mfrow=c(1,1))

plot(hsales.ts,ylab="House Sales",main = "Raw Data")

plot(diff(hsales.ts, lag=1, differences=1), main = "Lag-1 Difference", ylab="House Sales")

plot(diff(hsales.ts, lag=12, differences=1),main = "Seasonal Difference",ylab="House Sales")

plot(diff(diff(hsales.ts, lag=12, differences=1),lag=1, differences=1),ylab="House Sales",main = "Double Difference")

DOUBLEDIFF <- diff(diff(hsales.ts, lag=12, differences=1),lag=1, differences=1)

kpss.test(DOUBLEDIFF) ## The data is now stationary,Null Hypothesis- stationary Time series

par(mfrow=c(1,1))
Acf(DOUBLEDIFF, lag.max= 20)
Pacf(DOUBLEDIFF, lag.max= 20)
```
## From above PACF plot we can estimate below parameters
## d = 1
## p = 0
## P = 2
## From above ACF plot we can estimate below parameters
## q = 2
## Q = 2

```{r}
###   (f) Fit an ARIMA model using auto.arima ()function in R. How good was your anticipated model?


## Manual Choice (ARIMA(0,1,2)(2,1,2))
automodel <- auto.arima(hsales.ts)
manualmodel <- Arima(hsales.ts, order = c(0,1,2), seasonal = c(2,1,2)) 
forecast.auto <- forecast(automodel,h=24)
forecast.manual <- forecast(manualmodel,h=24)

par(mfrow=c(1,1))
plot(forecast.auto)
lines(forecast.manual$mean, lwd=2, col="red")
legend(1996,2800, legend=c("Automatic", "Manual","Raw data"),  col=c("blue", "red","black"), lty=1:1, cex=1)

plot(forecast.auto$residuals,lwd=2, col="blue", ylab="Residuals")
lines(forecast.manual$residuals,lwd=2, col="red")
legend(1960,15, legend=c("Automatic", "Manual","Raw data"),  col=c("blue", "red","black"), lty=1:1, cex=1)


```


```{r}
automodel <- auto.arima(hsales.ts)
summary(automodel)
```
## Auto ARIMA does a multiple iterations for the plots and hence gives an optimized result for the best model
## Here seasonality is present in the residuals and hence [p,d,q] and [P,D,Q] orders are generated.
## d = 1 one times differencing is required.
## p = 0 no immediate spikes on PACF plot
## P = 2 2 significant spikes on PACF plot corresponding to seasons

## q = 2 2 significant spikes on ACF plot corresponding to MA-2
## Q = 0 no seasonal correlation for MA model





